Modelling the emergence of whisker barrels

### 2020-06-30 paper/briefpaper/brief.tex
Colour wrangling, label changing

% Blue changes are major point 2 (competition mechanism) and other, general
% uncategorised changes.
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
% Red for Stuart's first updates
\newcommand{\stu}[1]{\textcolor{red}{#1}}
% Minor point 4:
\newcommand{\mpfour}[1]{\textcolor{revgreen}{#1}}
% Minor point 6:
\newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
% Justify competition:
\newcommand{\mpfive}[1]{\textcolor{revpurp2}{#1}}
% Justify gamma/interaction parameter choice. major point 1.
\newcommand{\MPone}[1]{\textcolor{revpink}{#1}}
% Describe the metrics
\newcommand{\metrics}[1]{\textcolor{revred}{#1}}
% Parameter search. Major point 3
\newcommand{\paramsearch}[1]{\textcolor{revbblack}{#1}}
% Sensitivity analysis. Major point 3
\newcommand{\sensitivity}[1]{\textcolor{revgblack}{#1}}

@@

\cite{karbowski_model_2004} developed a reaction-diffusion style model of how
extrinsic signalling gradients can constrain the emergence of distinct fields
from intrinsic cortical dynamics. Their model defines how the density of
connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a
1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
$i$. The model was derived from the assumption that the rates at which $a_i$
and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
$c_i(\mb{x},t)$, and model synaptogenesis as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
\end{equation}
%
Accordingly, where the total density of synaptic connections sums to one,
connections decay at rate $\alpha$. Otherwise the connection density increases
non-linearly ($k>1$) with the density of axon branching. Axon branching is
modelled as
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \cmnt{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
\end{equation}
%
The first term on the right describes the divergence (indicated by
$\nabla\vcdot$) of the quantity in parentheses, which is referred to as the
`flux' of axonal branching. The flux represents diffusion across the cortical
sheet, at rate $D$, and the influence of $M$ molecular signalling fields,
$\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given
thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which
may be positive or negative in order that axons may branch in the direction of
either higher or lower concentrations. Note that computing the divergence in
simulation requires cells on the cortical sheet to communicate with immedately
adjacent cells only (see \emph{Materials \& Methods}). The second term on the
right quantifies the coupling between axon branching and synaptogenesis. Here
  $\chi_i=0$ is a placeholder.

@@

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}, \cmnt{Fig.\,5b}). \textbf{C} Simulation results for parameters
      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \cmnt{the}
      simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
      line); black squares \cmnt{provide a} measure \cmnt{of} the correspondence
      between the real and
      simulated barrel shapes, $\eta(t)$ (the \cmnt{pattern quality metric---see
        Materials \& Methods; units mm$^3$});
      \cmnt{grey hexagons \cmnt{provide a measure of how selectively each
          cortical site is innervated};
        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
      \textbf{E} \cmnt{Plotted across the cortical sheet, the selectivity
        develops to reveal an alignment with the emergent barrel boundary
        shapes}. \cmnt{The colour map
        indicates the values of $\mu(\mb{x})$. Overlaid
        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

@@

\mpfive{Testing this reasoning required increasing the strength of the
  competition between simulated thalamocortical projections for cortical
  territory, by increasing the tendency for the density of each projection to
  concentrate. The major modification required was thus to introduce into the
  model an additional source of competition between thalamic projections.}
%
The term in parentheses in Eq.\,\ref{eq:dc} represents competition between
thalamocortical projections for a limited availability of cortical
connections. To introduce competition also in terms of axon branching,
\cmnt{whilst ensuring that $a_i$ is conserved over time,} we
redefined
%
\begin{equation} \label{eq:comp}
  \color{blue}
  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
  \color{black}
\end{equation}
%
\mpfive{This term contributes to the \emph{flux} of axonal branching as
  an additional source of diffusion, which reduces the branching density for a
  given projection where the branches of other projections are dense. The
  strength of this additional competition is parameterized by
  $\epsilon$}. Note that this operation is local to individual afferent
projections.

@@

From initial conditions sample from a uniform random distribution for
$a(\mb{x},0)\in(0.2,0.4)$ and $c(\mb{x},0)=0$, and across a wide range of
parameter values,
%From random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
%  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
%  $c(\mb{x},0)=0$},
a clear Voronoi-like tesselation \cmnt{of topographically organised
  projections} emerged, \cmnt{confirming that barrel maps can self-organise in
  the absence of pre-specified center points. Parameters for the example
  simulation shown in Fig.\,\ref{fig:main}C (see Movie S1) were obtained by
  conducting a full parameter sweep and choosing a combination of values that
  scored well against three measures.}

@@

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the
        model given in Eqs.\,\ref{eq:dc}--\ref{eq:comp}. Graphs show the three
        pattern metrics for $t=30000$ steps as colourmaps: Honda $\delta$ (top
        row) shows conformance to a Voronoi pattern; $\eta$ (middle row) gives
        the simulated/experimental barrel area difference and $\omega$ (bottom
        row) is a measure of selectivity of the connection densities for the
        41 thalamocortical axon types. The colour maps are chosen so that for
        each row, lighter (orange and yellow) colours indicate desirable
        properties for the simulations: Lower values of $\delta$ are yellow,
        indicating good Voroni patterns; Lower values of $\eta$ indicate good
        pattern matching; Higher values of $\omega$ indicate greater
        selectivity. White squares indicate that the simulation was not stable
        for that parameter set.
%
        The parameter space explored is three dimensional with the parameter
        ratio $\alpha/\beta$ varied along the columns of graphs, the
        competition parameter $\epsilon$ varied on the $y$-axes and the
        diffusion constant, $D$ varied along the $x$-axes. An asterisk (*)
        marks the parameter set displayed in Fig.\,\ref{fig:main}. Boxes i),
        ii)~and iii)~mark parameter sets for which corresponding patterns are
        shown in \textbf{B} (for $t=30000$ steps).
%
        \textbf{B} The effect of varying the diffusion constant,
        $D$. i)~Increasing $D$ above the optimum value gives an expansion of
        the outer barrels and a corresponding compression of the barrels
        within. ii)~Conversely, an expansion of the central barrels and
        compression of those near the boundary results from reducing $D$ from
        its optimum. iii) Reducing the diffusion constant is equivalent to
        increasing the size of the domain $S$ and so this simulation, in which
        $D$ had the lowest stable value, can be thought to represent an animal
        with a large cortical area but with unchanged cortical complexity. A
        large proportion of the domain is occupied by the projections with
        intermediate interaction parameters; those with strong interaction
        parameters are compressed around the edge of the domain. A barrel
        pattern fails to form.}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\paramsearch{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
  iterations, plotted as colour maps, for 216 parameter sets. The maps
  indicate that the pattern is not strongly influenced by the choice of
  $\alpha/\beta$, which was varied by 5 orders of magnitude. The most
  Voronoi-like patterns were achieved for parameter sets for which $D$ and
  $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the solution of
  maximum selectivity, which favours lower values of $D$. Thus, there is a
  trade-off between good selection and good Voroni pattern conformation.  The
  parameter set indicated by the asterisk in Fig.\,\ref{fig:paramsweep}A was
  chosen for Fig.\,\ref{fig:main} because it provides good pattern
  correspondence and a close approximation to a Voronoi tesselation.}

\paramsearch{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Because the patterns are robust
  to alterations in $\epsilon$ and the choices of $\alpha$ and $\beta$, we
  concentrate our attention on the diffusion constant, $D$. Changing $D$,
  whose value defines the rate at which a region of localized branching will
  spread out across the domain, $S$, is equivalent to making changes to the
  size of $S$ with a fixed branching diffusion rate. These simulations can
  therefore be thought of as representing animals with a varying cortical
  areas but with unchanged cortical complexity. The most extreme example
  (Fig.\,\ref{fig:paramsweep}B iii) could represent an animal with a very
  large cortical area (such as the South American capybara) which receives
  sensory input from $N=41$ thalamic barreloids. Here, a large proportion of
  the domain is occupied by the projections with intermediate interaction
  parameters and those with strong interaction parameters are bunched up
  towards the edge of the domain. In this case, a barrel pattern fails to
  form, agreeing with the observation that cortical whisker barrels are absent
  in the capybara.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig3.png}
    \caption{
      \sensitivity{\textbf{A} Sensitivity of the system pattern
        metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
        (bottom row), evaluated at $t=30000$ steps,
        to three classes of perturbation. $y$-axes and colour scales have identical ranges
        to the colour scales in Fig.\,\ref{fig:paramsweep}, to ease comparison
        with the parameter exploration.
        %
        \emph{Column 1:} Random noise drawn from a uniform distribution,
        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
        to the values for the interaction parameters, $\gamma_{i,j}$ used in
        Fig.\,\ref{fig:main}.
        %
        \emph{Column 2:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
        was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
        was smoothed by convolution with a symmetric 2D
        Gaussian kernel of width $\sigma_\rho$. Colour maps in this column
        thus show the effect of changing the magnitude and the length scale of
        noise in the guidance fields.
        %
        \emph{Column 3:} Colour maps show the result of setting the rotational angle
        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
        parameters of $\rho_2(\mb{x})$
        unchanged from those used in the simulation in
        Fig.\,\ref{fig:main}, for which
        $\phi_{\rho_2}=84$\textdegree and $G_{\rho_2}=1$.
        %
        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
        the patterns. i) The effect of significant interaction parameter noise
        causes barrels to become swapped from the positions they would
        otherwise take. ii) High magnitude, short lengthscale noise in
        $\rho(\mb{x})$ leads to `wiggly' boundaries between barrels. iii)
        Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)
        which reduces the strength of the axons' interaction with the guidance
        field in the left-right sense. Barrel rows B, C and D become `crushed'
        down the center line, and edge barrels dominate. iv) The effect of
        rotating $\rho_1$ by 20\textdegree~is a relatively mild distortion of
        the pattern in the anticlockwise sense.
      }
    }
    \label{fig:sens}
  \end{fullwidth}
\end{figure}

% This section needs to emphasize that the absolute values of gamma are not important
\sensitivity{We performed a sensitivity analysis to determine if the quality
  of the pattern (i.e.~the values of $\delta$, $\eta$ \& $\omega$ after
  $t=30000$ steps) was affected by perturbations to any of the following
  inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
}

\sensitivity{First, we conducted 10 runs of the simulation shown in
  Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
  metrics in the default model. The results (where deviations occur as a
  result of the different random initial conditions for $a_i$) were
  $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
  $\omega=0.2165\,(0.0001)$. The pattern was barely affected by the random
  changes to $a_i(t=0)$ on each run.}
%
\sensitivity{Next, we modified the parameters of the generating uniform
  distribution to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$,
  increasing the variability of $a_i(\mb{x},t=0)$ without changing the mean
  initial activation. We ran this system 10 times and obtained the results
  $\delta=0.090\,(0.005)$, $\eta=0.2136\,(0.004)$ and
  $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests demonstrated that the
  increase in noise amplitude made no significant difference to the metric
  results ($\delta: \mathrm{t}=-0.69,\mathrm{p}=0.49$, $\eta:
  \mathrm{t}=-2.0,\mathrm{p}=0.06$, $\omega:
  \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a noise-free initial condition
  with $a_i=0.3$ for every hex showed that the model is deterministic (all
  runs produced the results $\delta=0.0928$, $\eta=0.21236$ and
  $\omega=0.21649$) and that stochasticity in $a_i(\mb{x},t=0)$ is not
  required by this model, as all three results lay within one standard
  deviation of those for the runs in which $a_i$ was initialized
  stochastically.}

@@

\sensitivity{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to
  iv). The results show that the Voronoi nature of the pattern ($\delta$) and
  the selectivity of the fields are not strongly affected by interaction
  parameter noise, but the pattern metric, $\eta$ shows a degradation in the
  pattern arrangement when a) noise in the interaction parameters causes their
  disordering (find better way to express?), b) noise in the guidance field
  with a high magnitude and short length scale causes disruption to the barrel
  edges and c) the guidance fields tend towards a non-orthogonal arrangement,
  or if one field becomes substantially stronger in its interaction with the
  thalamocortical projections. Fig\,\ref{fig:sens}B shows perturbed examples.}

@@

\stu{In reality, scores of smaller Dirichletform barrels representing the
  microvibrissae form to the right of where the E-row barrels emerge in our
  simulations, presumably via the same competitive processes. Enforcing here
  the same boundary condition as used to represent the true edges of the
  barrel field is necessary to ensure the stability of the simulation, though
  we acknowledge that this region of the boundary is enforced primarily to
  keep the number of simulated projections, and hence the overall
  computational complexity of the simulation, manageable (simulating each
  additional projection introduces an extra \textbf{YYY} new dynamical
  variables).}

>>>>

% Miscellaneous changes
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
% Justify gamma/interaction parameter choice. major point 1.
\newcommand{\MPone}[1]{\textcolor{revpink}{#1}}
% Blue changes are major point 2 (competition mechanism) and other, general
% uncategorised changes.
\newcommand{\MPtwo}[1]{\textcolor{revblue}{#1}}
% Parameter search. Major point 3
\newcommand{\MPthreePar}[1]{\textcolor{revbblack}{#1}}
% Sensitivity analysis. Major point 3
\newcommand{\MPthreeSens}[1]{\textcolor{revgblack}{#1}}
% Justify competition:
\newcommand{\mpfive}[1]{\textcolor{revpurp2}{#1}}
% Minor point 4:
\newcommand{\mpfour}[1]{\textcolor{revgreen}{#1}}
% Minor point 6:
\newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
% Describe the metrics
\newcommand{\metrics}[1]{\textcolor{revred}{#1}}

@@

\cite{karbowski_model_2004} developed a reaction-diffusion style model of how
extrinsic signalling gradients can constrain the emergence of distinct fields
from intrinsic cortical dynamics. Their model defines how the density of
connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a
1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
$i$. The model was derived from the assumption that the rates at which $a_i$
and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
$c_i(\mb{x},t)$, and model synaptogenesis as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
\end{equation}
%
Accordingly, where the total density of synaptic connections sums to one,
connections decay at rate $\alpha$. Otherwise the connection density increases
non-linearly ($k>1$) with the density of axon branching. Axon branching is
modelled as
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \MPtwo{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
\end{equation}
%
The first term on the right describes the divergence (indicated by
$\nabla\vcdot$) of the quantity in parentheses, which is referred to as the
`flux' of axonal branching. The flux represents diffusion across the cortical
sheet, at rate $D$, and the influence of $M$ molecular signalling fields,
$\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given
thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which
may be positive or negative in order that axons may branch in the direction of
either higher or lower concentrations. Note that computing the divergence in
simulation requires cells on the cortical sheet to communicate with immedately
adjacent cells only (see \emph{Materials \& Methods}). The second term on the
right quantifies the coupling between axon branching and synaptogenesis. Here
  $\chi_i=0$ is a placeholder.

@@

\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig1.png}
    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
      by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
      measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
      perfect Voronoi tesselation. Right shows the initial distribution of axon
      branching density ($a$) for one thalamocortical projection, and two
      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
        traced from \textbf{A}}. \textbf{B} The strengths of interaction
      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
      projections by the lengths of green and blue arrows respectively, assuming
      that similar fields aligned to the posterior-anterior and medial-lateral
      axes in the ventroposterior medial nucleus of the thalamus are sampled at
      the locations of putative barreloid centers (reconstructed from
      \citealp{haidarliu_size_2001}, \MPone{Fig.\,5b}). \textbf{C} Simulation results for parameters
      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
      projection for which the connection density is maximal and black lines
      delineate boundaries (see Movie
      S1). \textbf{D} Red dots show the Honda metric, \metrics{$\delta(t)$}, obtained from \metrics{the}
      simulation approaching that obtained from \metrics{real} barrels in \textbf{A} (dotted
      line); black squares \metrics{provide a} measure \metrics{of} the correspondence
      between the real and
      simulated barrel shapes, $\eta(t)$ (the \metrics{pattern quality metric---see
        Materials \& Methods; units mm$^3$});
      \metrics{grey hexagons \metrics{provide a measure of how selectively each
          cortical site is innervated};
        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
      \textbf{E} \metrics{Plotted across the cortical sheet, the selectivity
        develops to reveal an alignment with the emergent barrel boundary
        shapes}. \metrics{The colour map
        indicates the values of $\mu(\mb{x})$. Overlaid
        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
    \label{fig:main}
  \end{fullwidth}
\end{figure}

@@

\mpfive{Testing this reasoning required increasing the strength of the
  competition between simulated thalamocortical projections for cortical
  territory, by increasing the tendency for the density of each projection to
  concentrate. The major modification required was thus to introduce into the
  model an additional source of competition between thalamic projections.}
%
The term in parentheses in Eq.\,\ref{eq:dc} represents competition between
thalamocortical projections for a limited availability of cortical
connections. To introduce competition also in terms of axon branching,
\MPtwo{whilst ensuring that $a_i$ is conserved over time,} we
redefined
%
\begin{equation} \label{eq:comp}
  \color{revblue}
  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
  \color{black}
\end{equation}
%
\mpfive{This term contributes to the \emph{flux} of axonal branching as
  an additional source of diffusion, which reduces the branching density for a
  given projection where the branches of other projections are dense. The
  strength of this additional competition is parameterized by
  $\epsilon$}. Note that this operation is local to individual afferent
projections.

@@

From initial conditions sample from a uniform random distribution for
$a(\mb{x},0)\in(0.2,0.4)$ and $c(\mb{x},0)=0$, and across a wide range of
parameter values,
%From random initial conditions for $a(\mb{x},0)$, \MPtwo{with each hexagonal element
%  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
%  $c(\mb{x},0)=0$},
a clear Voronoi-like tesselation \cmnt{of topographically organised
  projections} emerged, \cmnt{confirming that barrel maps can self-organise in
  the absence of pre-specified center points. Parameters for the example
  simulation shown in Fig.\,\ref{fig:main}C (see Movie S1) were obtained by
  conducting a full parameter sweep and choosing a combination of values that
  scored well against three measures.}

@@

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\MPthreePar{\textbf{A} Exploring the parameter space of the
        model given in Eqs.\,\ref{eq:dc}--\ref{eq:comp}. Graphs show the three
        pattern metrics for $t=30000$ steps as colourmaps: Honda $\delta$ (top
        row) shows conformance to a Voronoi pattern; $\eta$ (middle row) gives
        the simulated/experimental barrel area difference and $\omega$ (bottom
        row) is a measure of selectivity of the connection densities for the
        41 thalamocortical axon types. The colour maps are chosen so that for
        each row, lighter (orange and yellow) colours indicate desirable
        properties for the simulations: Lower values of $\delta$ are yellow,
        indicating good Voroni patterns; Lower values of $\eta$ indicate good
        pattern matching; Higher values of $\omega$ indicate greater
        selectivity. White squares indicate that the simulation was not stable
        for that parameter set.
%
        The parameter space explored is three dimensional with the parameter
        ratio $\alpha/\beta$ varied along the columns of graphs, the
        competition parameter $\epsilon$ varied on the $y$-axes and the
        diffusion constant, $D$ varied along the $x$-axes. An asterisk (*)
        marks the parameter set displayed in Fig.\,\ref{fig:main}. Boxes i),
        ii)~and iii)~mark parameter sets for which corresponding patterns are
        shown in \textbf{B} (for $t=30000$ steps).
%
        \textbf{B} The effect of varying the diffusion constant,
        $D$. i)~Increasing $D$ above the optimum value gives an expansion of
        the outer barrels and a corresponding compression of the barrels
        within. ii)~Conversely, an expansion of the central barrels and
        compression of those near the boundary results from reducing $D$ from
        its optimum. iii) Reducing the diffusion constant is equivalent to
        increasing the size of the domain $S$ and so this simulation, in which
        $D$ had the lowest stable value, can be thought to represent an animal
        with a large cortical area but with unchanged cortical complexity. A
        large proportion of the domain is occupied by the projections with
        intermediate interaction parameters; those with strong interaction
        parameters are compressed around the edge of the domain. A barrel
        pattern fails to form.}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\MPthreePar{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
  iterations, plotted as colour maps, for 216 parameter sets. The maps
  indicate that the pattern is not strongly influenced by the choice of
  $\alpha/\beta$, which was varied by 5 orders of magnitude. The most
  Voronoi-like patterns were achieved for parameter sets for which $D$ and
  $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the solution of
  maximum selectivity, which favours lower values of $D$. Thus, there is a
  trade-off between good selection and good Voroni pattern conformation.  The
  parameter set indicated by the asterisk in Fig.\,\ref{fig:paramsweep}A was
  chosen for Fig.\,\ref{fig:main} because it provides good pattern
  correspondence and a close approximation to a Voronoi tesselation.}

\MPthreePar8u{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Because the patterns are robust
  to alterations in $\epsilon$ and the choices of $\alpha$ and $\beta$, we
  concentrate our attention on the diffusion constant, $D$. Changing $D$,
  whose value defines the rate at which a region of localized branching will
  spread out across the domain, $S$, is equivalent to making changes to the
  size of $S$ with a fixed branching diffusion rate. These simulations can
  therefore be thought of as representing animals with a varying cortical
  areas but with unchanged cortical complexity. The most extreme example
  (Fig.\,\ref{fig:paramsweep}B iii) could represent an animal with a very
  large cortical area (such as the South American capybara) which receives
  sensory input from $N=41$ thalamic barreloids. Here, a large proportion of
  the domain is occupied by the projections with intermediate interaction
  parameters and those with strong interaction parameters are bunched up
  towards the edge of the domain. In this case, a barrel pattern fails to
  form, agreeing with the observation that cortical whisker barrels are absent
  in the capybara.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig3.png}
    \caption{
      \MPthreeSens{\textbf{A} Sensitivity of the system pattern
        metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
        (bottom row), evaluated at $t=30000$ steps,
        to three classes of perturbation. $y$-axes and colour scales have identical ranges
        to the colour scales in Fig.\,\ref{fig:paramsweep}, to ease comparison
        with the parameter exploration.
        %
        \emph{Column 1:} Random noise drawn from a uniform distribution,
        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
        to the values for the interaction parameters, $\gamma_{i,j}$ used in
        Fig.\,\ref{fig:main}.
        %
        \emph{Column 2:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
        was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
        was smoothed by convolution with a symmetric 2D
        Gaussian kernel of width $\sigma_\rho$. Colour maps in this column
        thus show the effect of changing the magnitude and the length scale of
        noise in the guidance fields.
        %
        \emph{Column 3:} Colour maps show the result of setting the rotational angle
        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
        parameters of $\rho_2(\mb{x})$
        unchanged from those used in the simulation in
        Fig.\,\ref{fig:main}, for which
        $\phi_{\rho_2}=84$\textdegree and $G_{\rho_2}=1$.
        %
        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
        the patterns. i) The effect of significant interaction parameter noise
        causes barrels to become swapped from the positions they would
        otherwise take. ii) High magnitude, short lengthscale noise in
        $\rho(\mb{x})$ leads to `wiggly' boundaries between barrels. iii)
        Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)
        which reduces the strength of the axons' interaction with the guidance
        field in the left-right sense. Barrel rows B, C and D become `crushed'
        down the center line, and edge barrels dominate. iv) The effect of
        rotating $\rho_1$ by 20\textdegree~is a relatively mild distortion of
        the pattern in the anticlockwise sense.
      }
    }
    \label{fig:sens}
  \end{fullwidth}
\end{figure}

% This section needs to emphasize that the absolute values of gamma are not important
\MPthreeSens{We performed a sensitivity analysis to determine if the quality
  of the pattern (i.e.~the values of $\delta$, $\eta$ \& $\omega$ after
  $t=30000$ steps) was affected by perturbations to any of the following
  inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
}

\MPthreeSens{First, we conducted 10 runs of the simulation shown in
  Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
  metrics in the default model. The results (where deviations occur as a
  result of the different random initial conditions for $a_i$) were
  $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
  $\omega=0.2165\,(0.0001)$. The pattern was barely affected by the random
  changes to $a_i(t=0)$ on each run.}
%
\MPthreeSens{Next, we modified the parameters of the generating uniform
  distribution to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$,
  increasing the variability of $a_i(\mb{x},t=0)$ without changing the mean
  initial activation. We ran this system 10 times and obtained the results
  $\delta=0.090\,(0.005)$, $\eta=0.2136\,(0.004)$ and
  $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests demonstrated that the
  increase in noise amplitude made no significant difference to the metric
  results ($\delta: \mathrm{t}=-0.69,\mathrm{p}=0.49$, $\eta:
  \mathrm{t}=-2.0,\mathrm{p}=0.06$, $\omega:
  \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a noise-free initial condition
  with $a_i=0.3$ for every hex showed that the model is deterministic (all
  runs produced the results $\delta=0.0928$, $\eta=0.21236$ and
  $\omega=0.21649$) and that stochasticity in $a_i(\mb{x},t=0)$ is not
  required by this model, as all three results lay within one standard
  deviation of those for the runs in which $a_i$ was initialized
  stochastically.}

@@

\MPthreeSens{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to
  iv). The results show that the Voronoi nature of the pattern ($\delta$) and
  the selectivity of the fields are not strongly affected by interaction
  parameter noise, but the pattern metric, $\eta$ shows a degradation in the
  pattern arrangement when a) noise in the interaction parameters causes their
  disordering (find better way to express?), b) noise in the guidance field
  with a high magnitude and short length scale causes disruption to the barrel
  edges and c) the guidance fields tend towards a non-orthogonal arrangement,
  or if one field becomes substantially stronger in its interaction with the
  thalamocortical projections. Fig\,\ref{fig:sens}B shows perturbed examples.}

@@

\cmnt{In reality, scores of smaller Dirichletform barrels representing the
  microvibrissae form to the right of where the E-row barrels emerge in our
  simulations, presumably via the same competitive processes. Enforcing here
  the same boundary condition as used to represent the true edges of the
  barrel field is necessary to ensure the stability of the simulation, though
  we acknowledge that this region of the boundary is enforced primarily to
  keep the number of simulated projections, and hence the overall
  computational complexity of the simulation, manageable (simulating each
  additional projection introduces an extra \textbf{YYY} new dynamical
  variables).}
